### 2022-06-20 article_v2/tex/SM.tex
article v2

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/SM_figure.pdf}
\end{figure}
\section{The Standard Model. Figure adapted from a diagram made by Carsten Burgard.}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight independent generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

@@

\subsection{Beyond the Standard Model}
Together, these three interactions make up the Standard Model gauge group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This determines the form of the three coupling constants, which numerical values must be experimentally measured. 
Since the vector bosons are represented by the generators, they are uniquely determined by the symmetry group. However, the scalar boson(s) and fermions are free as long as they
belong to representations of the symmetry group. By this construction, modifications to fermions rather than bosons are generally easier to make, allowing us to propose amendments to the model. Even the
number of fermions must be experimentally verified and can be altered from a phenomenological standpoint. In this work,
we will use this leniency to introduce a new particle and examine to which extent 
these modifications might be supported by experimental evidence. Moreover, we will examine the possibility of adding a new completely new interaction, which manifests itself as modifications to the neutrino-matter interactions.

\section{Mass Generation}
In the following sections, the presentation is following the outline in~\cite{giunti}.
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermion's right and left-handed components.
All neutrinos are left-handed, and all antineutrinos are right-handed~\cite{giunti}. This observational fact is reflected in the 
left-handed lepton doublet and right-handed lepton singlet

\begin{align}
    L^\prime_{\alpha L} = \begin{pmatrix}
            \nu^\prime_{\alpha L} \\
        \alpha^\prime_{L} 
    \end{pmatrix}\,, \quad
    \ell^\prime_{\alpha R} =
    \alpha^\prime_{R} \,,            
\end{align}
where $\alpha \in \{e,\mu,\tau\}$, and the primes denotes that the fields do not have definite masses.
The essential part here is that $\nu^\prime_{\alpha R}$ is missing from the right-handed lepton singlet.
Thus, the neutrino fields do not pick up a Yukawa coupling, 
and will not undergo the Higgs mechanism, which leaves them massless.
Moreover, all terms of a Lagrangian that we construct must respect the gauge invariance. This removes the possibility of the neutrino mass to be 
generated at loop level because any such attempt will violate the total lepton number by two units.
In order to keep our theory renormalizable, we have one option left if we want to generate neutrino masses: introducing right-handed neutrino and left-handed antineutrino fields.

We consider a right-handed neutrino field, $\nu^\prime_R$. Since the electroweak gauge group $\text{SU}(2)_L \times U(1)_Y$ only couple to 
left-handed particles and right-handed antiparticles, $\nu_R$ transforms as a singlet under the Standard Model symmetry 
group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$.

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa part of the interaction Lagrangian 
\begin{align} 
    \mathcal{L}_{int} \subseteq -\left( \frac{v + H}{\sqrt{2}} \right) \left[ \bar{\ell}^{\prime}_{\alpha L} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \bar{\nu}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,,
\end{align}
where $v$ is the Higgs vacuum expectation value, $H$ the Higgs field, $\ell_{\alpha L}^\prime (\ell_{\beta R}^\prime)$ the left (right) handed charged lepton field, and summation over the lepton indices $\alpha,\beta$ is implied. $Y_{\alpha \beta}^{\prime \ell}$ are the charged lepton Yukawa couplings, and $\nu^\prime_{\alpha L}(\nu^\prime_{\beta R})$ are the left (right) handed neutrino fields .
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, as emphasized with their prime. 
This can be remedied by diagonalizing $Y^{\prime \ell}$  with a biunitary transformation
\begin{align}
    V_{ L}^{\ell \dagger} Y^{\prime \ell} V_{ R}^{\ell}=Y^{\ell} \,.
\end{align}
where $V^\ell_L$ and $V^\ell_R$ are two unitary matrices.
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. 
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then -- by definition -- the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

So, let the neutrino field in the mass basis have components with Latin numerals to distinguish them from the flavour components, i.e 
\begin{align}\label{eq:nu_rotation} 
    \nu_{k} =  V_{k\beta}^{\nu \dagger} \nu_{\beta}^\prime\,.
\end{align}
The diagonalized Lagrangian now takes the form 
\begin{align}
    \mathcal{L}_{int} &\subseteq -\left( \frac{v + H}{\sqrt{2}} \right) \left[\bar{\ell}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \bar{\nu}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right] \nonumber \\
    &= -\left( \frac{v + H}{\sqrt{2}} \right) \left[\bar{\ell}_{\alpha L}^{\prime} V_{\alpha \beta L}^{\ell} Y_{\alpha \beta}^{ \ell} V_{\alpha \beta R}^{\ell \dagger} \ell_{\beta R}^{\prime}
    + \bar{\nu}_{\alpha L}^{\prime} V_{\alpha k L}^{\nu} Y_{kj}^{\nu} V_{\beta j  R}^{\nu \dagger} \nu_{\beta R}^{\prime}\right] \nonumber \\
    &= -\left( \frac{v + H}{\sqrt{2}} \right) \left[\bar{\ell}_{\alpha L} Y_{\alpha \beta}^{ \ell} \ell_{\beta R} + \bar{\nu}_{k L} Y_{kj}^{ \nu} \nu_{j R}\right] \nonumber \\
    &= -\left( \frac{v + H}{\sqrt{2}} \right) \left[\bar{\ell}_{\alpha L} Y_{\alpha \beta}^{ \ell} \ell_{\beta R} + \bar{\nu}_{k L} Y_{kj}^\nu \nu_{j R}\right]
\end{align}
By construction, $Y^\ell$ and $Y^\nu$ are diagonal, so we write their components as $y_{\alpha}^{\ell} \delta_{\alpha \beta}$ and $y_{k}^{\nu} \delta_{k j}$ respectively, leaving the Lagrangian as 
\begin{align}\label{eq:L_H}
    \mathcal{L}_{int} 
    &\subseteq-\left( \frac{v + H}{\sqrt{2}} \right) \left[\bar{\ell}_{\alpha L} y_{\alpha}^{\ell} \delta_{\alpha \beta} \ell_{\beta R} + \bar{\nu}_{k L} y_{k}^{\nu} \delta_{k j} \nu_{j R}\right] \nonumber \\
    &=-\left( \frac{v + H}{\sqrt{2}} \right) \left[\bar{\ell}_{\alpha L} y_{\alpha}^{\ell}  \ell_{\alpha R} + \bar{\nu}_{k L} y_{k}^{\nu} \nu_{k R}\right] \nonumber \\
    &=-\left( \frac{v + H}{\sqrt{2}} \right) \left[ y_{\alpha}^{\ell}  \bar{\ell}_{\alpha L}\ell_{\alpha R} +  y_{k}^{\nu}\bar{\nu}_{k L} \nu_{k R}\right] 
\end{align}
Now, by the introduction of the right-handed field $\nu_R$, the Dirac neutrino field is
\begin{align}
    \nu_k = \nu_{kL} + \nu_{kR}\,.
\end{align}
Multiplying $\nu_k$ with its conjugate $\bar{\nu}_k$, we get 
\begin{align}
    \bar{\nu}_k \nu_k 
    & = \bar{\nu}_{k L} \nu_{k L} +\bar{\nu}_{k R}\nu_{k L} + \bar{\nu}_{k L}\nu_{k R} + \bar{\nu}_{k R}\nu_{k R} \nonumber \\
    & = \bar{\nu}_{k L}\nu_{k R} + \bar{\nu}_{k R}\nu_{k L} \nonumber \\
    & = \bar{\nu}_{k L}\nu_{k R} + \t{h.c.}
\end{align}
The same calculation for the charged lepton field yields the same result for $\ell_k$. Substituting this result and multiplying the Higgs vacuum expectation value $v$ into the fields gives us
\begin{align}
    \mathcal{L}_{H} 
    &=-\left( \frac{v + H}{\sqrt{2}} \right) \left[ y_{\alpha}^{\ell}   \bar{\ell}_\alpha \ell_\alpha  +  y_{k}^{\nu} \bar{\nu}_k \nu_k \right] \nonumber \\
    &=- \frac{y_{\alpha}^{\ell} v}{\sqrt{2}}   \bar{\ell}_\alpha \ell_\alpha   -  \frac{ y_{k}^{\nu} v}{\sqrt{2}} \bar{\nu}_k \nu_k  - \frac{y_{\alpha}^{\ell}}{\sqrt{2}}   \bar{\ell}_\alpha \ell_\alpha H  -  \frac{ y_{k}^{\nu}}{\sqrt{2}} \bar{\nu}_k \nu_k H\,.
\end{align}
Thus, this extension to the SM generates by the Higgs mechanism neutrino masses with terms
\begin{align}
    m_k = \frac{y_k^\nu v}{\sqrt{2}}\,,
\end{align}
where the Yukawa couplings $y^\nu_k$ needs to be experimentally determined.

>>>>

\section{The Standard Model}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight independent generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

@@

\subsection{Beyond the Standard Model}
Together, these three interactions make up the Standard Model gauge group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This determines the form of the three coupling constants, which numerical values must be experimentally measured. 
Since the vector bosons are represented by the generators, they are uniquely determined by the symmetry group. However, the scalar boson(s) and fermions are free as long as they
belong to representations of the symmetry group. By this construction, modifications to fermions rather than bosons are generally easier to make, allowing us to propose amendments to the model. 
We will use this leniency to will examine the possibility of adding a new completely new interaction, which manifests itself as modifications to the neutrino-matter interactions.

### 2022-06-20 article_v2/tex/abstract.tex
article v2

\chapter*{Abstract}
\addcontentsline{toc}{section}{Abstract}

The failure of the Standard Model of particle physics to predict neutrino masses invites us to amend it. We have no reason to believe that 
the interactions currently described by the gauge group are the whole picture, nor should we expect that the number of fermions hitherto observed is correct.
In this thesis, we explore two amendments to the Standard Model, and examine whether either of these are consistent with measured data.

Firstly, we consider the sterile neutrino, which has no interactions described by the Standard Model. We examine the effect of this 
new fermion on the neutrino oscillation probabilities, and present what could be a detectable signal in the \si{\TeV} energy range.
Secondly, we consider interactions beyond the Standard Model, possibly stemming from a higher-order theory.
We show how the parameters of these Non-Standard Interactions (NSI) can modify the oscillation probabilities and within which energy range we expect to discern this signal.
We use data from and simulate event counts in two Cherenkov detectors: IceCube and DeepCore. Moreover, we generate data and simulate a proposed upgrade of the DeepCore detector: PINGU.

Using IceCube track events, we obtain best-fit values $\dm[41] = \SI{0.01}{\eV^2}$ and $\theta_{24} = 0.67$ for our sterile neutrino hypothesis at 
a p-value of $20\%$, which is not statistically significant. Hence, we found no evidence of a sterile neutrino in IceCube data. 
Moreover, we were unable to distinguish a signal from $\theta_{34}$ in our IceCube simulation.
We obtain stringent bounds on the NSI parameters and compare those to previous results in literature. We show that PINGU is expected to narrow the bound further 
on $\emt$, especially by considering a joint analysis with IceCube and DeepCore. Finally, we see that an anti-correlation between 
$\eem$ and $\eet$ at probability level was propagated down to event level, which we expect to be observable by PINGU.

\chapter*{Sammanfattning}
\addcontentsline{toc}{section}{Sammanfattning}
Misslyckandet med partikelfysikens så kallade standardmodell att förutsäga neutrinomassor uppmanar oss att förbättra den. Vi har ingen anledning att tro att
de interaktioner som för närvarande beskrivs av standardmodellens symmetrigrupp fullständigt beskriver universum, och inte heller vi bör förvänta oss att antalet fermioner som hittils observerats
är korrekt.
I detta examensarbete undersöker vi två förlängningar av standardmodellen och undersöker om dessa är förenliga med uppmätta data.
Först introducerar vi den sterila neutrinon, som inte har några interaktioner som beskrivs av standardmodellen. Vi undersöker effekten av denna
nya fermion på neutrinoscillationssannolikheterna, och presentera vad som skulle vara en detekterbar signal i \si{\TeV}-området.
Vidare introducerar vi en ny interaktion som vi kallar Icke-Standard Interaktioner (NSI), som möjligen härrör från en högre ordningens teori.
Vi visar hur NSI-parametrarna kan modifiera oscillationssannolikheterna, och inom vilka energiområden vi förväntar oss att urskilja denna signal.
Vi simulerar två Cherenkov-detektorer: IceCube och DeepCore. Dessutom presenterar vi resultat som vi
kan förvänta oss att se från en föreslagen detektor: PINGU.

Genom att analysera spår från muoner i IceCube erhåller vi bästa anpassningsvärden $\dm[41] = \SI{0.01}{\eV^2}$ och $\theta_{24} = 0.67$ för vår sterila neutrinohypotes, med
ett p-värde på $20\%$, vilket inte är statistiskt signifikant. Därmed hittade vi inga bevis på en steril neutrino i IceCube-datat.
Dessutom kunde vi inte urskilja en signal från $\theta_{34}$ i vår IceCube-simulering.
Vi uppmäter snäva gränser för NSI-parametrarna och jämför dem med tidigare resultat i litteraturen. Vi visar att PINGU förväntas minska gränsen ytterligare
hos $\emt$, särskilt ifall man utformar en gemensam analys med IceCube och DeepCore. Slutligen ser vi att en anti-korrelation mellan
$\eem$ och $\eet$ på sannolikhetsnivå spreds ner till händelsenivå, något som vi förväntar oss kan observeras av PINGU.

\chapter*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}
First and foremost, I would like to thank my supervisor Prof. Sandhya Choubey for her extensive guidance and support throughout,
and for the time she dedicated to my project.

Secondly, I'm eternally grateful to everyone who contributed to making me be the person who I am. 
My parents Anders and Helena, and also Ingemar. Thank you for nurturing my curiosity and for your unwavering belief in me.
My friends from KTH,
mainly Svante, Anders, and Erik for making the final years -- in which I got to know you -- the best. 
My best friends, Gilbert and Max, with whom I have shared countless of evenings filled with laughter and 
great conversations.

Finally, beloved girlfriend Moa, who I met at the start of this thesis. I have no idea if this would have been possible without you.

\vspace{9cm}
The computations were enabled by resources provided by the Swedish
National Infrastructure for Computing (SNIC) at HPC2N, Umeå University
and NSC, Linköping University. The resources were partially funded by the Swedish Research 
Council through grant agreement no. 2018-05973

>>>>

\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

We consider interactions beyond the Standard Model, possibly stemming from a higher-order theory.
We show how the parameters of these Non-Standard Interactions (NSI) can modify the oscillation probabilities and within which energy range we expect to discern this signal.
We use data from and simulate event counts in two Cherenkov detectors: IceCube and DeepCore. Moreover, we generate data and simulate a proposed upgrade of the DeepCore detector: PINGU.

We obtain stringent bounds on the NSI parameters and compare those to previous results in literature. We show that PINGU is expected to narrow the bound further 
on $\emt$, especially by considering a joint analysis with IceCube and DeepCore. Finally, we see that an anti-correlation between 
$\eem$ and $\eet$ at probability level was propagated down to event level, which we expect to be observable by PINGU.

\section*{Acknowledgements}
The computations were enabled by resources provided by the Swedish
National Infrastructure for Computing (SNIC) at HPC2N, Umeå University
and NSC, Linköping University. The resources were partially funded by the Swedish Research 
Council through grant agreement no. 2018-05973

### 2022-06-20 article_v2/tex/conclusion.tex
article v2

\chapter{Conclusion}\label{ch:conc}
In this thesis, we have explored the potential of explaining new physics using detectors handled by the IceCube collaboration.
We started off by introducing the Standard Model in Sec.~\ref{ch:SM}, and noted that it has several shortcomings. 
We mentioned the Super-Kamiokande experiment and that their 1998 result found no other explanation than that $\nm$ oscillated~\cite{sk1998}, in direct violation of the Standard Model. 
We saw that, by adding a right-handed neutrino field, neutrino masses are generated by the Higgs mechanism, thus enabling neutrino oscillations to occur. 
In Sec.~\ref{ch:oscillation}, we saw how the neutrino mixing matrix could be constructed and derived formulas describing neutrino oscillations in the two- and three-neutrino pictures. Adding a term for matter potential,
we then arrived at a Hamiltonian that we used in the time-dependent Schrödinger equation to solve for the oscillation probabilities in matter.

In Sec.~\ref{sec:anomalies}, we learned that the LSND experiment found an excess in neutrino events that could be explained by a fourth, sterile, neutrino $\ns$~\cite{lsnd}.
We modified the Hamiltonian to incorporate mixing with this new neutrino species, and saw how a \si{\eV^2} scale  $\ns$ with $\theta_{24}\neq 0$ affects the $\Pamam$ oscillation probability 
by giving us a resonant \si{\TeV} $\anm$ disappearance.

We then put the sterile neutrino aside and introduced Non-Standard Interactions (NSI) in Sec.\ref{sec:nsiTheory} as a possible example of new physics beyond the Standard Model.
We then studied the effects of NSI on the neutrino oscillation probabilities and saw that NSI effects other than those stemming from $\emt$ are not apparent in the \si{\TeV} region.
Moreover, we saw in $\Pme$ that we could expect a correlation between $\eem$ and $\eet$ in the single-digit \si{\GeV} range.
We also saw that $\emt$ effects were visible on probability level in both the \si{\GeV} and \si{\TeV} regions.

First, we described our method in testing the sterile neutrino hypothesis (3+1). We simulated the 86 string IceCube detector with track events, explained in Sec.\ref{ch:ICmethod}.
After simulating event counts in Sec.~\ref{sec:sterileResults} for the $3+1$ hypothesis, we were able to find best-fit parameters of 
$\dm[41] = \SI{0.01}{\eV^2}$ and $\theta_{24} = 0.67$ ($\sin^22\theta_{24} = 0.95$) at 
a p-value of $20\%$. Even though we found a best-fit for two parameters of the sterile neutrino, they are not statistically significant. Hence, we found no evidence of a sterile neutrino in IceCube data.
We also investigated the case when $\theta_{24} = \theta_{34}$, which produced an identical contour as in the case with $\theta_{34}=0$. From this observation, 
we concluded that IceCube is currently insensitive to any impact of $\theta_{34}$, with track events alone. However,
when opening up the $\tau$ channel by allowing $\theta_{34} \neq 0$, it is possible that with the inclusion of cascade events from $\nt$, IceCube could be able to better observe 
an impact of $\theta_{34}$. With these results, we then completely left the sterile neutrino behind.

In Sec.~\ref{sec:constraining}, we moved on to the NSI parameter $\emt$ and its effect events captured by IceCube. 
Using a $\chi^2$ test, we found that the bound from IceCube on $\emt$ at 90 \% confidence level was 
\begin{align}
    -0.015 <\,& \emt < 0.014\ \quad (90\% \text{ CL})
\end{align}
for our `baseline' case of 15\% uncorrelated systematic uncertainty.

@@

In Sec.~\ref{ch:PINGUmethod} we described our method of simulating a proposed upgrade to the IceCube array: PINGU.
Since PINGU is not yet live, we produced data for it assuming a standard $3+0$ hypothesis with parameters from~\cite{nufit}.
After a $\chi^2$ analysis, we obtained the following constrains on the NSI parameters at 90\% confidence level:
\begin{align}
    -0.049 <&\, \ett < 0.057 \nonumber \\
    -0.010 <&\, \emt < 0.011 \nonumber \\
    -0.137 <&\, \eem < 0.11 \nonumber \\
    -0.132 <&\, \eet < 0.125 \quad (\text{all at }90\% \text{ CL})\,.
 \end{align}
After that, we did a joint $\chi^2$ analysis combining IceCube, DeepCore and simulated PINGU events to further constrict $\emt$. The bound at 90\% confidence level obtained was
\begin{align}
    -0.010 <&\, \emt < 0.0060\,. \quad (90\% \text{ CL})
 \end{align}

>>>>

\section*{Conclusion}\label{ch:conc}
We introduced Non-Standard Interactions (NSI) in Sec.~\ref{sec:nsiTheory} as a possible example of new physics beyond the Standard Model.
We then studied the effects of NSI on the neutrino oscillation probabilities and saw that NSI effects other than those stemming from $\emt$ are not apparent in the \si{\TeV} region.
Moreover, we saw in $\Pme$ that we could expect a correlation between $\eem$ and $\eet$ in the single-digit \si{\GeV} range.
We also saw that $\emt$ effects were visible on probability level in both the \si{\GeV} and \si{\TeV} regions.

We simulated the 86 string IceCube detector with track events, explained in Sec.~\ref{ch:ICmethod}.
In Sec.~\ref{sec:constraining}, we studied the NSI parameter $\emt$ and its effect events captured by IceCube. 
Using a $\chi^2$ test, we found that the bound from IceCube on $\emt$ at 90 \% confidence level was 
\begin{align}
    -0.015 <\,& \emt < 0.014\ \quad (90\% \text{ CL})
\end{align}
for our `baseline' case of 15\% uncorrelated systematic uncertainty.

@@

In Sec.~\ref{ch:PINGUmethod} we described our method of simulating a proposed upgrade to the IceCube array: PINGU.
Since PINGU is not yet live, we produced data for it assuming a standard $3$ neutrino hypothesis with parameters from~\cite{nufit}.
After a $\chi^2$ analysis, we obtained the following constrains on the NSI parameters at 90\% confidence level:
\begin{align}
    -0.049 <&\, \ett < 0.057 \nonumber \\
    -0.010 <&\, \emt < 0.011 \nonumber \\
    -0.137 <&\, \eem < 0.11 \nonumber \\
    -0.132 <&\, \eet < 0.125 \quad (\text{all at }90\% \text{ CL})\,.
 \end{align}
After that, we did a joint $\chi^2$ analysis combining IceCube, DeepCore and simulated PINGU events to further constrict $\emt$. The bound at 90\% confidence level obtained was
\begin{align}
    -0.010 <&\, \emt < 0.0060\,. \quad (90\% \text{ CL})
 \end{align}

### 2022-06-20 article_v2/tex/introduction.tex
article v2

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we introduce a new particle: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we set the sterile neutrino aside altogether and instead turn to the interactions between neutrinos and the Earth. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

Finally, Chapter~\ref{ch:results} contains the result from our two separate extensions of the Standard Model.
In Section~\ref{sec:sterileResults} we present our findings for the sterile neutrino and compare those with literature. 
In Section~\ref{sec:constraining}, we present new constrained bounds on the NSI parameters by combining our results for IceCube, DeepCore, and PINGU.

>>>>

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we consider a particle beyond the Standard Model: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we set the sterile neutrino aside altogether and instead turn to the interactions between neutrinos and mattter. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

Finally, Chapter~\ref{ch:results} contains the result from our two separate extensions of the Standard Model.
In Section~\ref{sec:sterileResults} we present our findings for the sterile neutrino. 
In Section~\ref{sec:constraining}, we present new constrained bounds on the NSI parameters compare those with literature.

### 2022-06-20 article_v2/tex/nsiResults.tex
article v2

\section{The NSI Hypothesis}\label{sec:constraining}
In this section, we will constrain the four NSI parameters $\ett$, $\emt$, $\eem$, and $\eet$. As we shall see, the impacts of NSI mainly manifesting 
themselves at \si{\GeV} energies, causing us to bring back DeepCore and PINGU. We will consider the results obtained from these three detectors separately as well as jointly.

@@

\subsection{Constraining the NSI parameters}\label{sec:constraining_NSI}
We let each of the four NSI parameters considered assume a value, while keeping the other three fixed at zero. We consider the following ranges of values for each parameter:
\begin{align}
   -0.07 \le &\, \ett \le 0.07 \nonumber \\
   -0.03 \le &\, \emt \le 0.03 \nonumber \\
   -0.3 \le &\, \eem \le 0.3 \nonumber \\
   -0.3 \le &\, \eet \le 0.3 \nonumber \\
\end{align}
After the oscillation parameters $\dm[31]$ and $\theta_{23}$ have been marginalized out, we plot $\Delta \chi^2$ for each of the four NSI parameters in Fig.~\ref{fig:3D_NO}. 
The results are shown in Fig.~\ref{fig:IC_3D} and summarized in Tables~\ref{table:IC_DC_results} and~\ref{table:PINGU_joint_results}. Each region is bounded
by the best and worst case scenario, as defined in Table~\ref{table:syst_errors}, while the middle line is the baseline scenario. We again emphasize that 
no uncertainty scenarios are considered for DeepCore, since they already provided in the data release~\cite{DC2019data}.

Comparing the PINGU and the DeepCore results in Fig.~\ref{fig:3D_NO}, we note that the best-fit for each NSI parameter for the PINGU experiment is expected to be zero. This is because the `data' we generated during 
the PINGU simulations assume no NSI since they have yet to be observed in nature. This introduces a non-NSI bias in all joint analyses which include PINGU
since PINGU has stronger statistics than DeepCore and will thus pull the joint $\chi^2$ towards $\epsilon =0$.
Moreover, we see that we can expect PINGU to be sensitive to systematic uncertainty, especially when constraining $\eem$ and $\eet$ from the negative side.

@@

We compare our results to a similar analysis performed on the same DeepCore dataset in~\cite{demidov}. In that publication, 
the constraints found at 90\% CL were 
\begin{align}\label{eq:demidov}
   -0.055 <&\, \ett < 0.056 \nonumber \\
   -0.023 <&\, \emt < 0.016 \nonumber \\
   -0.21 <&\, \eem < 0.20 \nonumber \\
   -0.19 <&\, \eet < 0.20\,.
\end{align}
Comparing~\ref{eq:demidov} to our results from Table~\ref{table:IC_DC_results} at 90\% CL, namely
\begin{align}\label{eq:DC_results}
   -0.054 <&\, \ett < 0.067 \nonumber \\
   -0.029 <&\, \emt < 0.0070 \nonumber \\
   -0.12 <&\, \eem < 0.15 \nonumber \\
   -0.084 <&\, \eet < 0.15\,,
\end{align}
we see that our results for $\eem$, $\eet$ are more stringent, while the results for $\emt$ and $\ett$ are similar but more lenient. It is worth noting that the 
results from~\cite{demidov} are have been obtained by including more systematic uncertainties, such as the spectral index, hadron production, and baryon resonance. 
However, the overall flux normalization in~\cite{demidov} is 20\%, while we opted for a slightly higher figure of 25\% which could compensate for the exclusion of the 
statistical uncertainties mentioned.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.

>>>>

\section{The NSI Hypothesis}\label{sec:constraining}
In this section, we will constrain the four NSI parameters $\ett$, $\emt$, $\eem$, and $\eet$. As we shall see, the impacts of NSI mainly manifest
themselves at \si{\GeV} energies, causing us to bring back DeepCore and PINGU. We will consider the results obtained from these three detectors separately as well as jointly.

@@

\subsection{Constraining the NSI parameters}\label{sec:constraining_NSI}
We let each of the four NSI parameters considered assume a value, while keeping the other three fixed at zero. We consider the following ranges of values for each parameter:
\begin{align}
   -0.07 \le &\, \ett \le 0.07 \nonumber \\
   -0.03 \le &\, \emt \le 0.03 \nonumber \\
   -0.3 \le &\, \eem \le 0.3 \nonumber \\
   -0.3 \le &\, \eet \le 0.3 \nonumber \\
\end{align}
After the oscillation parameters $\dm[31]$ and $\theta_{23}$ have been marginalized out, we plot $\Delta \chi^2$ for each of the four NSI parameters in Fig.~\ref{fig:3D_NO}. 
The results are shown in Fig.~\ref{fig:IC_3D} and summarized in Tables~\ref{table:IC_DC_results} and~\ref{table:PINGU_joint_results}. Each region is bounded
by the best and worst-case scenario, as defined in Table~\ref{table:syst_errors}, while the middle line is the baseline scenario. We again emphasize that 
no uncertainty scenarios are considered for DeepCore, since they already provided in the data release~\cite{DC2019data}.

Comparing the PINGU and the DeepCore results in Fig.~\ref{fig:3D_NO}, we note that the best-fit for each NSI parameter for the PINGU experiment is expected to be zero. This is because the `data' we generated during 
the PINGU simulations assume no NSI since they have yet to be observed in nature. This introduces a non-NSI bias in all joint analyses, which include PINGU
since PINGU has stronger statistics than DeepCore and will thus pull the joint $\chi^2$ towards $\epsilon =0$.
Moreover, we see that we can expect PINGU to be sensitive to systematic uncertainty, especially when constraining $\eem$ and $\eet$ from the negative side.

@@

We compare our results to a similar analysis performed on the same DeepCore dataset in~\cite{demidov}. In that publication, 
the constraints found at 90\% CL were 
\begin{align}\label{eq:demidov}
   -0.055 <&\, \ett < 0.056 \nonumber \\
   -0.023 <&\, \emt < 0.016 \nonumber \\
   -0.21 <&\, \eem < 0.20 \nonumber \\
   -0.19 <&\, \eet < 0.20\,.
\end{align}
Comparing~\ref{eq:demidov} to our results from Table~\ref{table:IC_DC_results} at 90\% CL, namely
\begin{align}\label{eq:DC_results}
   -0.054 <&\, \ett < 0.067 \nonumber \\
   -0.029 <&\, \emt < 0.0070 \nonumber \\
   -0.12 <&\, \eem < 0.15 \nonumber \\
   -0.084 <&\, \eet < 0.15\,,
\end{align}
we see that our results for $\eem$, $\eet$ are more stringent, while the results for $\emt$ and $\ett$ are similar but more lenient. It is worth noting that the 
results from~\cite{demidov} are have been obtained by including more systematic uncertainties, such as the spectral index, hadron production, and baryon resonance. 
However, the overall flux normalization in~\cite{demidov} is 20\%, while we opted for a slightly higher figure of 25\%, which could compensate for the exclusion of the 
statistical uncertainties mentioned.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector and illustrates the expected sensitivity of DeepCore for the NSI parameters.
